---
title: "R training"
author: "Karissa Willits"
date: 2026-03-18 
format: html
editor: visual
---

```{r}
source("training_source.R")
```


I am learning R! Don't need # when using chunks

```{r}
#use # when leaving notes in the code
# i dont know what ``` is for
#ctr enter just runs that line of code.
```


```{r}
3 #this will get just printed
4
typeof(8.5) #this is returning a function . returns the data type
#console important for tracking what happened. trying ot troubleshoot, look in the console. most of code writing want to be within a code file so its saved. 
```
```{r}
"text strings are fun"
typeof("strings ")
```
```{r}
TRUE
FALSE
typeof(TRUE)
```

```{r}
a <- c(1, 2, 3) #assign this as a
a
c("a", "b", "c")
c(TRUE, FALSE)
c(1, "two", 3)
#only returned characters. "" means it interpreted it as a character. Characters are always in quotes
#vector is one cell in Excel that has 1,2,3,4 all in the same cell. 


```

##lists (can hold mixed types)

```{r}
csv_path <- "../data/raw/presurvey_20260317.csv"
survey_raw <- rio ::import(csv_path) ##use rio :: to make sure it uses Rio when importing. 

```


