---
title: "R training"
author: "Karissa Willits"
date: 2026-03-18 
format: html
editor: visual
---

```{r}
source("training_source.R")
```

I am learning R! Don't need \# when using chunks

```{r}
#use # when leaving notes in the code
# i dont know what ``` is for
#ctr enter just runs that line of code.
```

```{r}
3 #this will get just printed
4
typeof(8.5) #this is returning a function . returns the data type
#console important for tracking what happened. trying ot troubleshoot, look in the console. most of code writing want to be within a code file so its saved. 
```

```{r}
"text strings are fun"
typeof("strings ")
```

```{r}
TRUE
FALSE
typeof(TRUE)
```

```{r}
a <- c(1, 2, 3) #assign this as a
a
c("a", "b", "c")
c(TRUE, FALSE)
c(1, "two", 3)
#only returned characters. "" means it interpreted it as a character. Characters are always in quotes
#vector is one cell in Excel that has 1,2,3,4 all in the same cell. 


```

##lists (can hold mixed types)

```{r}
csv_path <- "../data/raw/presurvey_20260317.csv"
survey_raw <- rio ::import(csv_path)  

```

```{r}
rm(survey_raw)
gc()

```


```{r}
survey_raw |>
glimpse()

survey <- survey_raw |>
clean_names() |>
rename(
os = what_operating_system_are_you_using_for_this_workshop,
prior_exp = which_of_the_following_are_true_regarding_your_experiences_with_r_prior_to_this_workshop,
r_hope = what_do_you_hope_to_use_r_for_in_the_future,
travel_mode = the_workshop_will_be_held_on_the_psu_campus_what_is_your_intended_mode_of_travel_to_the_workshop_check_all_that_apply,
travel_direct = do_you_intend_to_travel_directly_from_home_to_the_workshop,
distance = how_far_will_you_likely_travel,
commute_time = what_is_the_approximate_travel_time_for_your_morning_commute_to_psu,
drink = what_is_your_morning_drink_of_choice,
movie_quote = what_is_the_first_movie_quote_that_comes_to_mind,
pineapple_ok = pineapple_is_an_acceptable_pizza_topping
)

glimpse(survey)
```

```{r}

linux <- c("linux", "ubuntu", "debian", "fedora", "arch")

paste(linux, collapse = "|")

survey2 <- survey |> 
  mutate(commute_digits = str_extract(commute_time, "[0-9]+"),
        distance_digits = str_extract(distance, "[0-9]\\.+"),
        commute_min = as.numeric(commute_digits),
        distance_mi = as.numeric(distance_digits),
        travel_mph = if_else(!is.na(commute_min) & commute_min > 0,
                             distance_mi / commute_min * 60, NA_real_),
                                      
os = str_to_lower(os),
os_family = case_when(str_detect(os, "win") ~ "Windows", 
                      str_detect(os, "mac|apple") ~ "Mac",
                      str_detect(os, paste(linux, collapse = "|")) ~ "Linux",
                      TRUE ~ "Other"
                      )
                      )


glimpse(survey2)


survey3 <- survey2 |> 
  select(os, os_family, distance_mi, commute_min, travel_mph)

head(survey3)

```

```{r}

```

