---
title: "R training"
author: "Karissa Willits"
date: 2026-03-18 
format: html
editor: visual
---

```{r}
source("training_source.R")
library(textclean)
```


I am learning R! Don't need \# when using chunks

```{r}
#use # when leaving notes in the code
# i dont know what ``` is for
#ctr enter just runs that line of code.
```

```{r}
3 #this will get just printed
4
typeof(8.5) #this is returning a function . returns the data type
#console important for tracking what happened. trying ot troubleshoot, look in the console. most of code writing want to be within a code file so its saved. 
```

```{r}
"text strings are fun"
typeof("strings ")
```

```{r}
TRUE
FALSE
typeof(TRUE)
```

```{r}
a <- c(1, 2, 3) #assign this as a
a
c("a", "b", "c")
c(TRUE, FALSE)
c(1, "two", 3)
#only returned characters. "" means it interpreted it as a character. Characters are always in quotes
#vector is one cell in Excel that has 1,2,3,4 all in the same cell. 


```

##lists (can hold mixed types)

```{r}
csv_path <- "../data/raw/presurvey_20260317.csv"
survey_raw <- rio ::import(csv_path)  

```


```{r}
survey_raw |>
glimpse()

survey <- survey_raw |>
clean_names() |>
rename(
os = what_operating_system_are_you_using_for_this_workshop,
prior_exp = which_of_the_following_are_true_regarding_your_experiences_with_r_prior_to_this_workshop,
r_hope = what_do_you_hope_to_use_r_for_in_the_future,
travel_mode = the_workshop_will_be_held_on_the_psu_campus_what_is_your_intended_mode_of_travel_to_the_workshop_check_all_that_apply,
travel_direct = do_you_intend_to_travel_directly_from_home_to_the_workshop,
distance = how_far_will_you_likely_travel,
commute_time = what_is_the_approximate_travel_time_for_your_morning_commute_to_psu,
drink = what_is_your_morning_drink_of_choice,
movie_quote = what_is_the_first_movie_quote_that_comes_to_mind,
pineapple_ok = pineapple_is_an_acceptable_pizza_topping
)

glimpse(survey)
```

```{r}

linux <- c("linux", "ubuntu", "debian", "fedora", "arch")

linux_pat <- oaste(linux, collapse = "|")

survey2 <- survey |> 
  mutate(commute_digits = str_extract(commute_time, "[0-9]+"),
        distance_digits = str_extract(distance, "[0-9]\\.+"),
        commute_min = as.numeric(commute_digits),
        distance_mi = as.numeric(distance_digits),
        travel_mph = if_else(!is.na(commute_min) & commute_min > 0, 
                             distance_mi / commute_min * 60, 
                             NA_real_),
                                      
os = str_to_lower(os),
os_family = case_when(str_detect(os, "win") ~ "Windows", 
                      str_detect(os, "mac|apple") ~ "Mac",
                      str_detect(os, paste(linux, collapse = "|")) ~ "Linux",
                      TRUE ~ "Other"
                      )
                      )





survey3 <- survey2 |> 
  select(os, os_family, distance_mi, commute_min, travel_mph)

head(survey3)

```


```{r}
#do this when words and numbers are all over the place. 

#should generate long_survey. Open it and can see column with distance and time, column with corresponding answers. 


num_dict_words <- paste(textclean::replace_number(seq_len(100)), collapse = "|") #number length

text_num_tbl <-tibble(num = seq_len(100)) |> 
  mutate(
    word = textclean::replace_number(num),
    word_num = as.character(num)
    ) |> 
  select(word, word_num)

long_survey <- survey2 |>
  select(distance, commute_time) |> 
  mutate(rid = row_number()) |> 
  pivot_longer(cols = c(distance, commute_time), 
               names_to = "num_text_var", 
               values_to = "numeric_text") |>  
  mutate(numeric_text = str_to_lower(numeric_text)) |> #but have it all lowercase. 
  filter(nchar(numeric_text)>2) |>
  mutate(num_words = str_extract_all(numeric_text, num_dict_words, simplify = T), #mutate arguments separated by ,
         num_words = case_when(nchar(num_words[,1])<2 ~ NA_character_,
                               nchar(num_words[,2])>2 ~
                                 paste(num_words[,1], num_words[,2], sep = "-"),
                               TRUE ~ num_words[,1]) #have to end with TRUE
                    ) |> 
  
  left_join(select(text_num_tbl, word, word_num), by = c("num_words" = "word")) |> 
  mutate(num_raw = case_when(!is.na(word_num) ~ word_num, 
                             str_detect(numeric_text, "\\ban\\b") ~"1",  #\\b word\\b puts a boundary on an so it only searches for an
                             TRUE ~ numeric_text),
         num = parse_number(num_raw)
  )
           
#next thing would be to deal with units. Create variables that are minutes and hours
          
         
         

```


```{r}
  wide_survey <- long_survey |> 
    pivot_wider(
  id_cols = .rid,
  names_from = num_text_var,
  values_from = numeric_text, 
  names_repair = "minimal") |> 
  select(-.rid)
```
